---
title: "HW 1"
output:
  html_document: default
  pdf_document: default
editor_options:
  markdown:
    wrap: sentence
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE,  
                      warning = FALSE, message = FALSE, 
                      fig.align = "center",
                      R.options = list(max.print=100))

# Edit the file starting below
```

### Enter your name and EID here: Suren Samir Bhakta ssb2943

**You will submit this homework assignment as a pdf file on Gradescope.**

*For all questions, include the R commands/functions that you used to find your answer (show R chunk). Answers without supporting code will not receive credit. Write full sentences to describe your findings.*

------------------------------------------------------------------------

## Part 1: (11 pts)

The dataset `mtcars` was extracted from the 1974 Motor Trend US magazine, and comprises fuel consumption and other aspects of automobile design and performance for different cars (1973-74 models).
Look up the documentation for this data frame with a description of the variables by typing `?mtcars` **in the console** pane.

### Question 1: (2 pt)

Take a look at the first 6 rows of the dataset by using an `R` function in the code chunk below.
Do you know about any (or all) of these cars?

```{r}
# the head function displays the first six rows of the data set
x<-mtcars$cyl
x<-mtcars[mtcars$cyl == 8,]
mean(x$cyl)

```

**Based off the data given by the head function, I was able to see the name(s) of the first six cars. I, however, do not have any previous knowledge about any of these cars listed.**

------------------------------------------------------------------------

### Question 2: (2 pts)

How many rows and columns are there in this data frame in total?

```{r}
# nrow and ncol used to show number of rows and columns in mtcars.
nrow(mtcars)
ncol(mtcars)
```

**Within the data set "mtcars," there are 32 rows and 11 columns.**

------------------------------------------------------------------------

### Question 3: (1 pt)

Save `mtcars` in your environment and name it as your `eid`.
From now on, use this new object instead of the built-in dataset.

```{r}
# the "reverse-arrow" allows me to not only save the dataset "mtcars" into my environment, but also rename it to my liking.
ssb2943 <- mtcars
```

**The formerly known data set "mtcars" is now an object under the name "ssb2943," my UTEID.**

------------------------------------------------------------------------

### Question 4: (2 pts)

When is your birthday?
Using indexing, grab the value of `mpg` that corresponds to the day of your birthday (should be a number between 1 and 31).

```{r}
# picks the intersection of the seventeenth row and first column(mpg)
ssb2943[17,1]
```

**My birthday of October 17, 2002 was used to find the equivalent mpg. By using logical indexing with the day 17, I found the mpg, column 1, aligned with my birthday to be 14.7 Miles/(US) gallon.**

------------------------------------------------------------------------

### Question 5: (2 pts)

Using logical indexing, count the number of rows in the dataset where the variable `mpg` takes on values greater than 30.

```{r}
# "Nested Arguements." First find the mpgs greater than thirty, and then use nrow to count the rows that exceeded 30mpg.

nrow(ssb2943[ssb2943$mpg > 30,])
```

**Using logical indexing, I was able to find that there were 4 rows where mpg was over 30. To do this I set the argument of nrow as the rows where mpg were above thirty to count how many cars exceeded 30mpg.**

------------------------------------------------------------------------

### Question 6: (2 pts)

Let's create a new variable called `kpl` which converts the fuel efficiency `mpg` in kilometers per liter.
Knowing that 1 mpg corresponds to 0.425 kpl, complete the following code and calculate the max kpl:

```{r}
# Add a new variable to the dataset
ssb2943$kpl <- ssb2943$mpg*0.425
max(ssb2943$kpl)


```

**After creating a new variable, kpl, which converts miles per gallon to kilometers to liter, I found the max kpl of my dataset to be 14.4075 kilometers per liter.**

------------------------------------------------------------------------

## Part 2: (6 pts)

Let's quickly explore another built-in dataset: `airquality` which contains information about daily air quality measruements in New York, May to September 1973.

### Question 7: (2 pts)

Calculate the mean `Ozone` (in ppb).
Why does it make sense to get this answer?
*Hint: take a look at the column `Ozone` in the dataset.*

```{r}
# I used the View function to analyze the data then the mean function to calculate the mean Ozone rating.
mean(airquality$Ozone)
#View(airquality)
```

**After running the mean function for the Ozone quality of New York City, all was returned was NA. After using the View function, I had realized that the mean was not calculated because some days the Ozone quality was either not calculated or not reported, leaving the value as "N/A."** ------------------------------------------------------------------------

### Question 8: (2 pts)

Look at the documentation for the function `mean()` by running `?mean` **in the console**.
What argument should be used to find the mean value that we were not able to get in the previous question?
What type of values does that argument take?

**The argument that should be used is na.rm. The na.rm argument takes True or False, a Boolean, to determine whether NA values should be stripped from the dataset before computation.**

------------------------------------------------------------------------

------------------------------------------------------------------------

### Question 9: (2 pts)

Sometimes the R documentation does not feel complete.
We wish we had more information or more examples.
Find a post online (include the link) that can help you use that argument in the `mean()` function.
Then finally find the mean ozone!
[This site shows how to use the na.rm argument within the mean function](https://www.statology.org/na-rm/)

```{r}

mean(airquality$Ozone, na.rm = TRUE)
```

**After performing research, I had found a site that shows how to use the na.rm function. After successfully learning how to use the code, I found the mean Ozone of New York City to be 42.12931ppb.**

------------------------------------------------------------------------

## Formatting: (3 pts)

Knit your file!
Into pdf directly or into html.

Is it working?
If not, try to decipher the error message (look up the error message, consult websites such as [stackoverflow](https://stackoverflow.com/) or [crossvalidated](https://stats.stackexchange.com/).

Once it knits in html, click on `Open in Browser` at the top left of the window pops out.
Print your html file into pdf from your browser.
Any issue?
Ask your classmates or TA!

------------------------------------------------------------------------

```{r, echo=F}
## DO NOT DELETE THIS BLOCK!
Sys.info()
```
